Restrictive Clearance Comparison

### 2025-08-22 data-raw/ClearanceProject-2-RelativeLikelihood.Rmd
Updates project file organization. Moves common functions to a common, sourced file.

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ggvenn)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

@@

```{r data_loading}
this_pk <- readRDS("data-raw/plasma_CvTfits.rds")
load("data-raw/chem_lists.RData")
load("data-raw/plasma_CvTfits_data.RData")
load("data-raw/rat_pars_human.RData")
load("data-raw/rat_pars_rat.RData")
all_chem_properties <- readRDS("data-raw/RCdata/ctxR_3comp2_cheminfo.rds")
# Here we load Kd/binding affinity data
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")

@@

```{r venn_diagrams}
member_list <- list(cvt = cvt_plasma_chems,
                    httk_3comp2 = human_3compchems)

ggvenn(member_list, fill_color = c("white", "white"), auto_scale = FALSE)

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "httk_vs_CvT_VennDiagram_human.png"),
       bg = "white",
       height = 3, width = 5
)

```

```{r httk_pr_rmsle}

# RMSLE formula
# LE = log1p(Conc_est) - log1p(Conc)
# SLE = LE^2
# RMSLE = sqrt(mean(SLE, na.rm = TRUE))

@@

combined_RMSLE <- combined_preds %>%
  ungroup() %>%
  mutate(
    restrictive_SLE = (log1p(httk_Preds_restrictive) - log1p(Conc))^2,
    nonrestrictive_SLE = (log1p(httk_Preds_nonrestrictive) - log1p(Conc))^2,
    invivoPK_SLE =  (log1p(Conc_est) - log1p(Conc))^2 
  ) %>%
  group_by(Chemical, Species, Route) %>%
  summarize(RMSLE_restrictive = sqrt(mean(restrictive_SLE, na.rm = TRUE)),
            RMSLE_nonrestrictive = sqrt(mean(nonrestrictive_SLE, na.rm = TRUE)),
            RMSLE_invivoPKfit = sqrt(mean(invivoPK_SLE, na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
     between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive~ "More than"
  ))

combined_RMSLE %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("RMSLE_"),
               names_to = "Source",
               values_to = "RMSLE") %>%
  mutate(Source = factor(Source,
                         levels = c("RMSLE_nonrestrictive",
                                    "RMSLE_invivoPKfit",
                                    "RMSLE_restrictive")
  )) %>%
  ggplot(aes(x = Source, y = RMSLE, group = Chemical)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  theme_bw()

combined_RMSLE %>%
  ggplot(aes(x = Chemical, y = RMSLE_invivoPKfit)) +
  geom_point(color = "green4") +
  geom_point(aes(y = RMSLE_nonrestrictive), shape = 6, size = 0.5) +
  geom_point(aes(y = RMSLE_restrictive), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = RMSLE_nonrestrictive, ymax = RMSLE_restrictive)) +
  facet_grid(~factor(comparison_group,
                     levels = c("Less than", "Between", "More than")),
             scales = "free_x", space = "free_x") +
  labs(y = "RMSLE") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank(),
        axis.title = element_text(size = 14),
        strip.text = element_text(size = 12))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "RMSLE_comparison_fullChart.png"),
       bg = "white",
       height = 3.8, width = 10
)

# Number of chemicals we have RMSLE's for
combined_RMSLE %>% pull(Chemical) %>% unique() %>% length()

```

@@

```{r relative-likelihood}

# AIC and relative likelihood
# LL from pre-calculated httk predictions
this_pk <- this_pk %>% fit_sigma.pk(preds = rpr_httk_preds,
                                    pred_col = "httk_Preds_restrictive")
this_pk <- this_pk %>% fit_sigma.pk(preds = rpr_httk_preds,
                                    pred_col = "httk_Preds_nonrestrictive")

@@

httk_rel_ll <- httk_AIC %>%
  distinct(Chemical, Species, model, AIC) %>%
  pivot_wider(names_from = model, values_from = AIC) %>%
  filter(!is.na(model_3comp2_nonrestrictive)) %>%
  rowwise() %>% 
  mutate(minAIC = min(c_across(starts_with("model_")), 
                      na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x)/2) * 100, 4)
  })) %>%
  mutate(across(starts_with("model_"), \(x) {
    ifelse(x < 1E-6, 0, x)
  })) %>%
  mutate(ll_label = factor(
    case_when(model_flat > 95 ~ "Undetermined",
              model_3comp2_nonrestrictive > 50 &
                model_3comp2_restrictive > 50 ~ "Undetermined (non-flat)",
              model_3comp2_nonrestrictive > 50 ~ "Nonrestrictive",
              model_3comp2_restrictive > 50 ~ "Restrictive")
  ))

httk_rel_ll %>% 
  select(Chemical, Species, starts_with("model_"), minAIC) %>%
  mutate(Chemical = forcats::fct_reorder2(Chemical,
                                         model_flat,
                                         minAIC,
                                         mean)
         ) %>%
  pivot_longer(cols = starts_with("model_"),
               names_to = "model", values_to = "relative_ll") %>%
  mutate(
    source = ifelse(grepl("_3comp2_", model), "httk", "invivoPKfit")
  ) %>%
  ggplot(aes(x = Chemical, y = relative_ll, color = model, shape = source)) +
  geom_point(size = 0.75, alpha = 0.9) +
  scale_color_manual(
    values = c('#33a02c', '#cc88ff', '#7b3294', 'orange'),
    breaks = c('model_empirical',
               'model_3comp2_restrictive',
               'model_3comp2_nonrestrictive',
               'model_flat')
  ) +
  scale_shape_manual(values = c(2, 3)) +
  labs(
    title = "Relative likelihoods",
    subtitle = "(rat parameterized with human CLint and fup values)",
    y = "More Likely Model >>>",
    color = ''
  ) +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "AIC_comparisons.png"),
       bg = "white",
       height = 5, width = 10
)

this_rel_evaluated_df <- rph_httk_clearance_df %>%
  inner_join(httk_rel_ll %>% distinct(Chemical, Species, ll_label)) %>%
  inner_join(this_tkstats) %>% # Css
  inner_join(rph_param_df %>% 
               distinct(Chemical,
                        forced_human_values,
                        Fabsgut.restrictive,
                        Fabsgut.nonrestrictive,
                        CLint)) %>%
  mutate(Chemical = forcats::fct_reorder(Chemical, halflife_restrictive)) %>%
  ungroup() %>%
  distinct()

halflife_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, halflife_restrictive,
           halflife_nonrestrictive,
           ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = halflife.tkstats, 
                      ymin = halflife_nonrestrictive,
                      ymax = halflife_restrictive),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "halflife (hours)") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

ggsave(halflife_pl, filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "halflife_fullChart.png"),
       bg = "white",
       height = 3, width = 12
)

css_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css.tkstats, Css_httk_nr, Css_httk_r,
         ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = Css.tkstats, ymin = Css_httk_nr, ymax = Css_httk_r),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = 'Chemicals', y = "Css") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

cltot_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, CLtot_Fgutabs.tkstats,
           CLtot_restrictive, CLtot_nonrestrictive, 
           Fabsgut.restrictive, Fabsgut.nonrestrictive,
         ll_label) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = CLtot_Fgutabs.tkstats,
                      ymin = CLtot_restrictive/Fabsgut.restrictive,
                      ymax = CLtot_nonrestrictive/Fabsgut.nonrestrictive),
                  size = 0.1, shape = 23) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = 'Chemicals', y = "CLtot/Fabsgut") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        strip.text = element_blank(),
        strip.background = element_blank(),
        axis.text.y = element_text(color = "black", size = 10),
        axis.title = element_text(size = 10))

(halflife_pl + theme(axis.title.x = element_blank())) /
  (css_pl+ theme(axis.title.x = element_blank())) /
  cltot_pl

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "tkcomparisons_fullChart.png"),
       bg = "white",
       height = 4.5, width = 12
)

```

@@

combined_function_list <- union(all_chem_functions,
                                target_chem_functions) # 25 total

pre_enrichment_mat <- inner_join(
  functional_uses_all %>%
    filter(
      DTXSID %in% human_3compchems, !(DTXSID %in% target_chems_human), # all non-targets
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_nontargets"),
  functional_uses_all %>%
    filter(
      DTXSID %in% target_chems_human, 
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_targets")
) %>%
  janitor::clean_names() %>%
  mutate(n_total_targets = length(target_chems_human),
         n_total_nontargets = length(human_3compchems) - n_total_targets,
         n_nontargets_out = n_total_nontargets - n_nontargets,
         n_targets_out = n_total_targets - n_targets) %>%
  select(!c(n_total_nontargets, n_total_targets))

quick_enrich <- pre_enrichment_mat %>%
  rowwise() %>%
  mutate(odds_ratio = fisher.test(
    matrix(c(n_targets, n_nontargets,
             n_targets_out, n_nontargets_out),
           nrow = 2),
    alternative = "greater")$estimate,
    pval = fisher.test(
    matrix(c(n_targets, n_nontargets,
             n_targets_out, n_nontargets_out),
           nrow = 2),
    alternative = "greater")$p.value
  ) %>%
  ungroup() %>%
  mutate(padj = 0.05/n()) %>%
  arrange(pval)

# Makeshift upset plot
upset_pl <- this_rel_evaluated_df %>%
  distinct(Chemical, Species, Css_httk_r,
           halflife.tkstats, halflife_restrictive,
           halflife_nonrestrictive,
           ll_label) %>%
  left_join(functional_uses_all %>%
            distinct(DTXSID, `Harmonized Functional Use`) %>%
            rename(Chemical = DTXSID) %>%
            filter(`Harmonized Functional Use` %in% c(
              "Biocide",
              "Pharmaceutical (EPA)",
              "Degradant/impurity (EPA)"))
          ) %>%
  # filter(!is.na(`Harmonized Functional Use`)) %>%
  mutate(val = ifelse(is.na(`Harmonized Functional Use`), NA, "Present")) %>%
  distinct(Chemical, ll_label, `Harmonized Functional Use`, val) %>%
  ggplot(aes(x = Chemical, y = `Harmonized Functional Use`, color = val)) + 
  geom_point() +
  scale_color_manual(values = "black") +
  facet_grid(cols = vars(ll_label), scales = "free_x", space = "free_x") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        plot.background = element_rect(color = "white"),
        strip.text = element_blank(),
        legend.title = element_blank(), axis.title.y = element_blank())

@@

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "tkcomparisons_withUpset.png"),
       bg = "white",
       height = 4.5, width = 12
)

@@

# Enrichment plot
quick_enrich %>%
  mutate(
    harmonized_functional_use = forcats::fct_reorder(
      harmonized_functional_use, odds_ratio)
    ) %>%
  filter(padj_val <= 0.5) %>%
  ggplot() +
  geom_col(aes(x = odds_ratio, 
               y = harmonized_functional_use,
                fill = padj_val),
           color = "black", linewidth = 0.2,
           width = 0.75) +
  scale_fill_viridis_b(option = "viridis", direction = -1,
                       breaks = c(0, 0.05, 0.1, 0.5),
                       limits = c(0, 0.5)
                       ) +
  scale_x_continuous(limits = c(0, 12),
                     expand = expansion()) +
  labs(x = "Enrichment Score", y = "",
       title = "Enrichment of Functional Use Categories",
       fill = "adjusted p-value") +
  theme_bw() +
  theme(legend.title = element_text(hjust = 0.5),
        axis.text = element_text(color = "black"),
        axis.text.y = element_text(face = "bold", size = 10),
        plot.title = element_text(hjust = 0.5))

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "FunctionalUse_Enrichment.png"),
       bg = "white",
       height = 5, width = 6.5
)

```

@@

```{r selecting_chemical_params}

# Testing to make sure there are some chemicals that have different clearances
count(rph_httk_clearance_df, CLtot_restrictive == CLtot_nonrestrictive)
# The majority (140) do!

```

@@

```{r load_Kd}
red_assay_res %>%
  distinct(DTXSID, Affinity.Kd.Med, Affinity.Kd.Low, Affinity.Kd.High,
           Fit) %>%
  na.omit() %>%
  mutate(Target_Chemical = (DTXSID %in% target_chems_human)) %>%

>>>>

```{r library_loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ComplexHeatmap)
library(ggalluvial)
library(flextable)
library(reticulate)
library(patchwork)
source("data-raw/RC_comparison_bkgd.R") # httk & dawson_2021 loaded here
```

@@

```{r data_loading}
# TODO: Adjust this to load in new prediction files
# Adjust rest of document as well!
this_pk <- readRDS("data-raw/plasma_CvTfits.rds")
load("data-raw/chem_lists.RData")
load("data-raw/plasma_CvTfits_data.RData")
load("data-raw/rat_pars_human.RData")
load("data-raw/rat_pars_rat.RData")
all_chem_properties <- readRDS("data-raw/RCdata/ctxR_3comp2_cheminfo.rds")
# Here we load Kd/binding affinity data
red_assay_res <- readRDS("data-raw/2019-05-06-REDassay_HTTK.rds")

@@

```{r venn_diagrams}
member_list <- list_to_matrix(list(
  CvTdb = cvt_plasma_chems,
  httk_human = human_chems,
  httk_rat = rat_chems
)) |> make_comb_mat()

png(
  paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "httk_vs_CvT_UpsetPlot.png"
  ),
  bg = "white",
  height = 4, width = 4, units = "in",
  res = 200
)

ComplexHeatmap::UpSet(
  member_list,
  set_order = c("CvTdb", "httk_rat", "httk_human"),
  comb_order = order(comb_size(member_list), decreasing = TRUE),
  top_annotation = upset_top_annotation(
    show_annotation_name = FALSE,
    member_list,
    axis = FALSE,
    add_numbers = TRUE,
    numbers_rot = 0,
    numbers_gp = gpar(fontsize = 10)
  ),
  right_annotation = upset_right_annotation(
    show_annotation_name = FALSE,
    member_list,
    axis = FALSE,
    add_numbers = TRUE,
    numbers_rot = 0,
    numbers_gp = gpar(fontsize = 10)
  ),
  pt_size = unit(5, "mm")
)
dev.off()
```

```{r httk_pr_rmsle}
# RMSLE formula
# LE = log1p(Conc_est) - log1p(Conc)
# SLE = LE^2
# RMSLE = sqrt(mean(SLE, na.rm = TRUE))

@@

combined_RMSLE <- combined_preds %>%
  ungroup() %>%
  mutate(
    restrictive_SLE = (log1p(httk_Preds_restrictive) - log1p(Conc))^2,
    nonrestrictive_SLE = (log1p(httk_Preds_nonrestrictive) - log1p(Conc))^2,
    invivoPK_SLE = (log1p(Conc_est) - log1p(Conc))^2
  ) %>%
  group_by(Chemical, Species, Route) %>%
  summarize(
    RMSLE_restrictive = sqrt(mean(restrictive_SLE, na.rm = TRUE)),
    RMSLE_nonrestrictive = sqrt(mean(nonrestrictive_SLE, na.rm = TRUE)),
    RMSLE_invivoPKfit = sqrt(mean(invivoPK_SLE, na.rm = TRUE))
  ) %>%
  ungroup() %>%
  mutate(comparison_group = case_when(
    between(RMSLE_invivoPKfit, RMSLE_nonrestrictive, RMSLE_restrictive) |
      between(RMSLE_invivoPKfit, RMSLE_restrictive, RMSLE_nonrestrictive) ~ "Between",
    RMSLE_invivoPKfit < RMSLE_nonrestrictive &
      RMSLE_invivoPKfit < RMSLE_restrictive ~ "Less than",
    RMSLE_invivoPKfit > RMSLE_restrictive &
      RMSLE_invivoPKfit > RMSLE_nonrestrictive ~ "More than"
  ))

combined_RMSLE %>%
  ungroup() %>%
  pivot_longer(
    cols = starts_with("RMSLE_"),
    names_to = "Source",
    values_to = "RMSLE"
  ) %>%
  mutate(Source = factor(Source,
    levels = c(
      "RMSLE_nonrestrictive",
      "RMSLE_invivoPKfit",
      "RMSLE_restrictive"
    )
  )) %>%
  ggplot(aes(x = Source, y = RMSLE, group = Chemical)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  theme_bw()

combined_RMSLE %>%
  ggplot(aes(x = Chemical, y = RMSLE_invivoPKfit)) +
  geom_point(color = "green4") +
  geom_point(aes(y = RMSLE_nonrestrictive), shape = 6, size = 0.5) +
  geom_point(aes(y = RMSLE_restrictive), shape = 4, size = 0.5) +
  geom_linerange(aes(ymin = RMSLE_nonrestrictive, ymax = RMSLE_restrictive)) +
  facet_grid(
    ~ factor(comparison_group,
      levels = c("Less than", "Between", "More than")
    ),
    scales = "free_x", space = "free_x"
  ) +
  labs(y = "RMSLE") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 12)
  )

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "RMSLE_comparison_fullChart.png"
  ),
  bg = "white",
  height = 3.8, width = 10
)

# Number of chemicals we have RMSLE's for
combined_RMSLE %>%
  pull(Chemical) %>%
  unique() %>%
  length()
```

@@

```{r relative-likelihood}
# AIC and relative likelihood
# LL from pre-calculated httk predictions
this_pk <- this_pk %>% fit_sigma.pk(
  preds = rpr_httk_preds,
  pred_col = "httk_Preds_restrictive"
)
this_pk <- this_pk %>% fit_sigma.pk(
  preds = rpr_httk_preds,
  pred_col = "httk_Preds_nonrestrictive"
)

@@

httk_rel_ll <- httk_AIC %>%
  distinct(Chemical, Species, model, AIC) %>%
  pivot_wider(names_from = model, values_from = AIC) %>%
  filter(!is.na(model_3comp2_nonrestrictive)) %>%
  rowwise() %>%
  mutate(minAIC = min(c_across(starts_with("model_")),
    na.rm = TRUE
  )) %>%
  ungroup() %>%
  mutate(across(starts_with("model_"), \(x) {
    signif(exp((minAIC - x) / 2) * 100, 4)
  })) %>%
  mutate(across(starts_with("model_"), \(x) {
    ifelse(x < 1E-6, 0, x)
  })) %>%
  mutate(ll_label = factor(
    case_when(
      model_flat > 95 ~ "Undetermined",
      model_3comp2_nonrestrictive > 50 &
        model_3comp2_restrictive > 50 ~ "Undetermined (non-flat)",
      model_3comp2_nonrestrictive > 50 ~ "Nonrestrictive",
      model_3comp2_restrictive > 50 ~ "Restrictive"
    )
  ))

httk_rel_ll %>%
  select(Chemical, Species, starts_with("model_"), minAIC) %>%
  mutate(Chemical = forcats::fct_reorder2(
    Chemical,
    model_flat,
    minAIC,
    mean
  )) %>%
  pivot_longer(
    cols = starts_with("model_"),
    names_to = "model", values_to = "relative_ll"
  ) %>%
  mutate(
    source = ifelse(grepl("_3comp2_", model), "httk", "invivoPKfit")
  ) %>%
  ggplot(aes(x = Chemical, y = relative_ll, color = model, shape = source)) +
  geom_point(size = 0.75, alpha = 0.9) +
  scale_color_manual(
    values = c("#33a02c", "#cc88ff", "#7b3294", "orange"),
    breaks = c(
      "model_empirical",
      "model_3comp2_restrictive",
      "model_3comp2_nonrestrictive",
      "model_flat"
    )
  ) +
  scale_shape_manual(values = c(2, 3)) +
  labs(
    title = "Relative likelihoods",
    subtitle = "(rat parameterized with human CLint and fup values)",
    y = "More Likely Model >>>",
    color = ""
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5)
  )

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "AIC_comparisons.png"
  ),
  bg = "white",
  height = 5, width = 10
)

this_rel_evaluated_df <- rph_httk_clearance_df %>%
  inner_join(httk_rel_ll %>% distinct(Chemical, Species, ll_label)) %>%
  inner_join(this_tkstats) %>% # Css
  inner_join(rph_param_df %>%
    distinct(
      Chemical,
      forced_human_values,
      Fabsgut.restrictive,
      Fabsgut.nonrestrictive,
      CLint
    )) %>%
  mutate(Chemical = forcats::fct_reorder(Chemical, halflife_restrictive)) %>%
  ungroup() %>%
  distinct()

halflife_pl <- this_rel_evaluated_df %>%
  distinct(
    Chemical, Species, Css_httk_r,
    halflife.tkstats, halflife_restrictive,
    halflife_nonrestrictive,
    ll_label
  ) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(
    aes(
      y = halflife.tkstats,
      ymin = halflife_nonrestrictive,
      ymax = halflife_restrictive
    ),
    size = 0.1, shape = 23
  ) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "halflife (hours)") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    axis.text.y = element_text(color = "black", size = 10),
    axis.title = element_text(size = 10)
  )

ggsave(halflife_pl,
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "halflife_fullChart.png"
  ),
  bg = "white",
  height = 3, width = 12
)

css_pl <- this_rel_evaluated_df %>%
  distinct(
    Chemical, Species, Css.tkstats, Css_httk_nr, Css_httk_r,
    ll_label
  ) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(aes(y = Css.tkstats, ymin = Css_httk_nr, ymax = Css_httk_r),
    size = 0.1, shape = 23
  ) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "Css") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    axis.text.y = element_text(color = "black", size = 10),
    axis.title = element_text(size = 10)
  )

cltot_pl <- this_rel_evaluated_df %>%
  distinct(
    Chemical, Species, CLtot_Fgutabs.tkstats,
    CLtot_restrictive, CLtot_nonrestrictive,
    Fabsgut.restrictive, Fabsgut.nonrestrictive,
    ll_label
  ) %>%
  ggplot(aes(x = Chemical)) +
  geom_pointrange(
    aes(
      y = CLtot_Fgutabs.tkstats,
      ymin = CLtot_restrictive / Fabsgut.restrictive,
      ymax = CLtot_nonrestrictive / Fabsgut.nonrestrictive
    ),
    size = 0.1, shape = 23
  ) +
  facet_grid(cols = vars(ll_label), space = "free_x", scales = "free_x") +
  scale_y_log10() +
  labs(x = "Chemicals", y = "CLtot/Fabsgut") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    axis.text.y = element_text(color = "black", size = 10),
    axis.title = element_text(size = 10)
  )

(halflife_pl + theme(axis.title.x = element_blank())) /
  (css_pl + theme(axis.title.x = element_blank())) /
  cltot_pl

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "tkcomparisons_fullChart.png"
  ),
  bg = "white",
  height = 4.5, width = 12
)
```

@@

combined_function_list <- union(
  all_chem_functions,
  target_chem_functions
) # 25 total

pre_enrichment_mat <- inner_join(
  functional_uses_all %>%
    filter(
      DTXSID %in% human_3compchems, !(DTXSID %in% target_chems_human), # all non-targets
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_nontargets"),
  functional_uses_all %>%
    filter(
      DTXSID %in% target_chems_human,
      `Harmonized Functional Use` %in% combined_function_list
    ) %>% distinct(DTXSID, `Harmonized Functional Use`) %>%
    count(`Harmonized Functional Use`, name = "n_targets")
) %>%
  janitor::clean_names() %>%
  mutate(
    n_total_targets = length(target_chems_human),
    n_total_nontargets = length(human_3compchems) - n_total_targets,
    n_nontargets_out = n_total_nontargets - n_nontargets,
    n_targets_out = n_total_targets - n_targets
  ) %>%
  select(!c(n_total_nontargets, n_total_targets))

quick_enrich <- pre_enrichment_mat %>%
  rowwise() %>%
  mutate(
    odds_ratio = fisher.test(
      matrix(
        c(
          n_targets, n_nontargets,
          n_targets_out, n_nontargets_out
        ),
        nrow = 2
      ),
      alternative = "greater"
    )$estimate,
    pval = fisher.test(
      matrix(
        c(
          n_targets, n_nontargets,
          n_targets_out, n_nontargets_out
        ),
        nrow = 2
      ),
      alternative = "greater"
    )$p.value
  ) %>%
  ungroup() %>%
  mutate(padj = 0.05 / n()) %>%
  arrange(pval)

# Makeshift upset plot
upset_pl <- this_rel_evaluated_df %>%
  distinct(
    Chemical, Species, Css_httk_r,
    halflife.tkstats, halflife_restrictive,
    halflife_nonrestrictive,
    ll_label
  ) %>%
  left_join(functional_uses_all %>%
    distinct(DTXSID, `Harmonized Functional Use`) %>%
    rename(Chemical = DTXSID) %>%
    filter(`Harmonized Functional Use` %in% c(
      "Biocide",
      "Pharmaceutical (EPA)",
      "Degradant/impurity (EPA)"
    ))) %>%
  # filter(!is.na(`Harmonized Functional Use`)) %>%
  mutate(val = ifelse(is.na(`Harmonized Functional Use`), NA, "Present")) %>%
  distinct(Chemical, ll_label, `Harmonized Functional Use`, val) %>%
  ggplot(aes(x = Chemical, y = `Harmonized Functional Use`, color = val)) +
  geom_point() +
  scale_color_manual(values = "black") +
  facet_grid(cols = vars(ll_label), scales = "free_x", space = "free_x") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    plot.background = element_rect(color = "white"),
    strip.text = element_blank(),
    legend.title = element_blank(), axis.title.y = element_blank()
  )

@@

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "tkcomparisons_withUpset.png"
  ),
  bg = "white",
  height = 4.5, width = 12
)

@@

# Enrichment plot
quick_enrich %>%
  mutate(
    harmonized_functional_use = forcats::fct_reorder(
      harmonized_functional_use, odds_ratio
    )
  ) %>%
  filter(padj_val <= 0.5) %>%
  ggplot() +
  geom_col(
    aes(
      x = odds_ratio,
      y = harmonized_functional_use,
      fill = padj_val
    ),
    color = "black", linewidth = 0.2,
    width = 0.75
  ) +
  scale_fill_viridis_b(
    option = "viridis", direction = -1,
    breaks = c(0, 0.05, 0.1, 0.5),
    limits = c(0, 0.5)
  ) +
  scale_x_continuous(
    limits = c(0, 12),
    expand = expansion()
  ) +
  labs(
    x = "Enrichment Score", y = "",
    title = "Enrichment of Functional Use Categories",
    fill = "adjusted p-value"
  ) +
  theme_bw() +
  theme(
    legend.title = element_text(hjust = 0.5),
    axis.text = element_text(color = "black"),
    axis.text.y = element_text(face = "bold", size = 10),
    plot.title = element_text(hjust = 0.5)
  )

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "FunctionalUse_Enrichment.png"
  ),
  bg = "white",
  height = 5, width = 6.5
)
```

@@

```{r selecting_chemical_params}
# Testing to make sure there are some chemicals that have different clearances
count(rph_httk_clearance_df, CLtot_restrictive == CLtot_nonrestrictive)
# The majority (140) do!
```

@@

```{r load_Kd}
red_assay_res %>%
  distinct(
    DTXSID, Affinity.Kd.Med, Affinity.Kd.Low, Affinity.Kd.High,
    Fit
  ) %>%
  na.omit() %>%
  mutate(Target_Chemical = (DTXSID %in% target_chems_human)) %>%
  ggplot(
    aes(
      x = forcats::fct_reorder(DTXSID, Affinity.Kd.Low)
    )
  ) +
  geom_linerange(aes(
    ymin  = Affinity.Kd.Low,
    ymax = Affinity.Kd.High,
    color = Target_Chemical
  ), alpha = 0.4) +
  geom_point(aes(
    y = Affinity.Kd.Med, color = Target_Chemical
  ), size = 0.75) +
  # facet_grid(cols = vars(Fit), space = "free_x", scales = "free_x") +
  scale_y_log10(labels = scales::label_log(digits = 1)) +
  scale_color_manual(values = c("black", "green4")) +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    panel.grid = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.position = "bottom"
  ) +
  labs(title = "Kd estimates across chemicals", x = NULL)

ggsave(
  filename = paste0(
    Sys.getenv("FIG_DIR"),
    "RestrictiveClearance_Analysis/",
    "CURRENT_", "Kd_estimates_initialsummary.png"
  ),
  bg = "white",
  height = 4, width = 9
)
